---
title: "Clustering with Spotify Streaming History"
author: "Khoi Trinh"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Let's Explore Clustering

```{r libraries, warning=FALSE, message=FALSE}
library(cluster)
library(NbClust)
library(factoextra)
library(dplyr)
library(kmed)
```

## First, we need some data

### Data description

The data I chose is my own Spotify streaming history for the past year; you can find how to get your own Spotify data [here](https://www.makeuseof.com/tag/download-privacy-data-spotify/)

Then, follow [these instructions](https://github.com/vlad-ds/spoty-records) to obtain the song traits.

### Misc data processing:

Read in the data

```{r load data}
spotify = read.csv("final.csv")
# drop columns to help us find only unique songs
spotify= subset(spotify, select=-c(1,3,6,7))
# now get only unique songs
spotify=unique(spotify)
```

The data had almost 2500 observations, out of those, only the numeric data will be considered.

And out of the numeric columns, we will drop columns 3,5,12,13 as they are not song traits.

```{r convert to factor}
numericData = spotify %>% #Add data
  dplyr::select(where(is.numeric)) #finds where is.numeric is true

# drop the mentioned columns
numericData <- subset(numericData, select=-c(3,5,12,13))
```

## Time To Create Some Clusters

### One last data processing step

Scale the data, and use that to create our clusters. We need to scale the data as most of the traits are < 1; but tempo are not

So scaling is needed to not skewed the clusters.

```{r scale}
clusterData <- scale(numericData)
```

### K means

For this clustering method, the `kmeans()` function from the stats package is used. Let's start with 2 clusters. Normally, there are ways to determine an optimal number of clusters, but for the sake of simplicity, let's stick to 2, maybe we can change it later.

```{r k mean}
kmean <- kmeans(clusterData,centers = 2, nstart = 10)
```

### Clusters Analysis

First, let's see the size of the clusters

```{r size}
kmeansize <- kmean$size
kmeansize

cluster <- kmean$cluster
cluster
```

We have 2 clusters, with size 1925 and 224. 

Visually, the clusters look like so. We can see that the clusters have a little bit of overlap, but overall, it looks good.

```{r k means visual}
fviz_cluster(kmean, data = clusterData, geom = "point", ellipse.type = "convex", 
             ggtheme = theme_bw())
```

### Cluster Intepretation

Let's look at the center of each cluster.

```{r k mean clusters}
kmeaninfo <- data.frame(kmean$centers, kmean$size)
kmeaninfo
```

Here are the explanation of the traits, from [Spotify's API documentation](https://developer.spotify.com/documentation/web-api/reference/get-audio-features)

We can see that cluster 1 have songs that are mixed louder, have higher energy and liveness.

Cluster 2 have songs that are higher in valence, acousticness, danceability, but lower energy(?)

Let's increase the number of clusters, to see if we get any more clear separation.

```{r k mean redux}
kmean2 <- kmeans(clusterData,centers = 3, nstart = 10)

kmean2$size
```

```{r k means visual redux}
fviz_cluster(kmean2, data = clusterData, geom = "point", ellipse.type = "convex", 
             ggtheme = theme_bw())
```

```{r k mean clusters redux}
kmeaninfo2 <- data.frame(kmean2$centers, kmean2$size)
kmeaninfo2
```

We can do the same analysis with these 3 clusters.